Robust-Adaptive Control of Linear Systems:\\ beyond Quadratic Costs

### 2020-06-03 main.tex
C_N,delta is a confidence region on theta, not A(theta)

To address this issue, we turn to the framework of \emph{robust} decision-making: instead of merely considering a point estimate of the dynamics, for any $N\in\Natural$, we build an entire \emph{confidence region} $\cC_{[N],\delta}$, illustrated in \Cref{fig:estimation}, that contains the true dynamics with high probability:
\begin{align}
\probability{A(\theta)\in \cC_{[N],\delta}} \geq 1-\delta,
\label{eq:confidence}
\end{align}
where $\delta\in(0,1)$. In \Cref{sec:estimation}, having observed a history $\cD_{[N]} = \{(x_n, y_n,u_n)\}_{n\in[N]}$ of transitions, our first contribution extends the work of \citet{Abbasi2011} who provide a confidence ellipsoid for the least-square estimator to our setting of feature matrices, rather than feature vectors.

The \emph{robust control objective} $V^r$ \citep{Bental2009,Bertsimas2011,Gorissen2015} aims to maximise the worst-case outcome with respect to this confidence region $\cC_{[N],\delta}$:
\begin{equation}
\label{eq:robust-control}
\sup_{\bu\in{(\Real^q)}^\Natural} {V^r(\bu)}, \qquad \text{ where }\qquad {V^r(\bu)} \eqdef \inf_{\substack{A(\theta) \in \cC_{[N],\delta} \\ \bom\in[\underline{\bom},\overline{\bom}]^\Natural}} \left[\sum_{n=N+1}^\infty \gamma^n R(x_n(\bu,\bom))\right],
\end{equation}
$\gamma\in(0,1)$ is a discount factor, and $x_n(\bu,\bom)$ is the state reached at step $n$ under controls $\bu$ and perturbations $\bom$ within the given admissible bounds $[\underline\omega(t),\overline\omega(t)]$. Maximin problems such as $\eqref{eq:robust-control}$ are notoriously hard if the reward $R$ has a simple form. However, without a restriction on the shape of functions $R$, we cannot hope to derive an explicit solution.
In our second contribution, we propose a robust MPC algorithm for solving \eqref{eq:robust-control} numerically. In \Cref{sec:prediction}, we leverage recent results from the uncertain system simulation literature to derive an \emph{interval predictor} $[\underline{x}(t),\overline{x}(t)]$ for the system \eqref{eq:dynamics}, illustrated in \Cref{fig:prediction}. For any $N\in\Natural$, his predictor takes the information on the current state ${x}_N$, the confidence region $\cC_{[N],\delta}$, planned control sequence $\bu$ and admissible perturbation bounds $[\underline{\omega}(t),\overline{\omega}(t)]$; and must verify the \emph{inclusion property}:
\begin{equation}
\label{eq:inclusion-property}
\underline{x}(t)\leq x(t)\leq\overline{x}(t), \forall t\geq t_N.
\end{equation}

@@

\begin{figure}
	\begin{minipage}[b]{0.49\linewidth}
		\centering
		\includegraphics[trim={4cm 0 0 0}, clip, width=0.5\linewidth]{img/ellipsoid}
		\caption{The model estimation procedure, running on the obstacle avoidance problem of \Cref{sec:experiments}. The confidence region $C_{[N],\delta}$ shrinks with the number of samples $N$.}
		\label{fig:estimation}
	\end{minipage}
	\hfill
	\begin{minipage}[b]{0.49\linewidth}
		\centering
		\includegraphics[trim={4cm 0 0 0}, clip, width=0.8\linewidth]{img/obstacle_small}
		\caption{The state prediction procedure running on the obstacle avoidance problem of \Cref{sec:experiments}. At each time step (red to green), we bound the set of reachable states under model uncertainty \eqref{eq:confidence}}
		\label{fig:prediction}
	\end{minipage}%
\end{figure}

@@

\textbf{Linear dynamics and quadratic costs} ~~ Several approaches have been proposed for cumulative regret minimisation in the LQ problem. In the \emph{Optimism in the Face of Uncertainty} paradigm, the best possible dynamics within a high-confidence region is selected under a controllability constraint, to compute the corresponding optimal control in closed-form by solving a Riccati equation. The results of 	\citep{abbasi-yadkori11a,Ibrahimi2013,Faradonbeh2017} show that this procedure achieves a $\tilde{\cO}\left(N^{1/2}\right)$ regret. Posterior sampling algorithms \citep{Ouyang2017,abeille18a} select candidate dynamics randomly instead, and obtain the same result. Other works use noise injection for exploration such as \citep{Dean2017,Dean2018}. However, neither optimism nor random exploration fit a critical setting, where ensuring safety requires instead to consider pessimistic outcomes. The work of \citet{Dean2017} is close to our setting: after an offline estimation phase, they estimate a simple regret between a minimax controller and the optimal performance. Our work differs in that it addresses a generic shape cost. 
Another work of interest is \citep{Rosolia2019} where worst-case generic costs are considered. However, they assume the knowledge of the dynamics, and their rollout-based solution only produces inner-approximations and do not yield any guarantee. In this paper, interval prediction is used to produce oversets, while a near-optimal control is found using a tree-based planning procedure.

@@

We can enclose this confidence ellipsoid $\eqref{eq:confidence-ellipsoid}$ into a polytope $C_{[N],\delta}$. For simplicity, we present here a simple but coarse strategy: bound the ellipsoid by its enclosing axis-aligned hypercube. We obtain
\begin{align}
    \label{eq:polytope}
     \cC_{[N],\delta} = \left\{ A_0 +\sum_{i=1}^{2^d}\alpha_{i}\Delta A_{i}: \alpha\geq 0,  \sum_{i=1}^{2^d}\alpha_{i}=1\right\} \text{ where } A_0 = A + \theta_{N,\lambda}^\transp\phi,\; h_i\in\{-1,1\}^d,
\end{align}
$\Delta A_{i} = {h_i} \sqrt{\frac{\beta_N(\delta)}{\lambda_{\max}(G_{N,\lambda})}}$. A tighter polytope derivation is presented in the Supplementary Material. % at the price of an increased computational cost required by the diagonalisation of $G_{N,\lambda}$.

>>>>

To address this issue, we turn to the framework of \emph{robust} decision-making: instead of merely considering a point estimate of the dynamics, for any $N\in\Natural$, we build an entire \emph{confidence region} $\cC_{[N],\delta}\subset\Theta$, illustrated in \Cref{fig:estimation}, that contains the true dynamics parameter with high probability:
\begin{align}
\probability{\theta\in \cC_{[N],\delta}} \geq 1-\delta,
\label{eq:confidence}
\end{align}
where $\delta\in(0,1)$. In \Cref{sec:estimation}, having observed a history $\cD_{[N]} = \{(x_n, y_n,u_n)\}_{n\in[N]}$ of transitions, our first contribution extends the work of \citet{Abbasi2011} who provide a confidence ellipsoid for the least-square estimator to our setting of feature matrices, rather than feature vectors.

The \emph{robust control objective} $V^r$ \citep{Bental2009,Bertsimas2011,Gorissen2015} aims to maximise the worst-case outcome with respect to this confidence region $\cC_{[N],\delta}$:
\begin{equation}
\label{eq:robust-control}
\sup_{\bu\in{(\Real^q)}^\Natural} {V^r(\bu)}, \qquad \text{ where }\qquad {V^r(\bu)} \eqdef \inf_{\substack{\theta \in \cC_{[N],\delta} \\ \bom\in[\underline{\bom},\overline{\bom}]^\Natural}} \left[\sum_{n=N+1}^\infty \gamma^n R(x_n(\bu,\bom))\right],
\end{equation}
$\gamma\in(0,1)$ is a discount factor, and $x_n(\bu,\bom)$ is the state reached at step $n$ under controls $\bu$ and perturbations $\bom$ within the given admissible bounds $[\underline\omega(t),\overline\omega(t)]$. Maximin problems such as $\eqref{eq:robust-control}$ are notoriously hard if the reward $R$ has a simple form. However, without a restriction on the shape of functions $R$, we cannot hope to derive an explicit solution.
In our second contribution, we propose a robust MPC algorithm for solving \eqref{eq:robust-control} numerically. In \Cref{sec:prediction}, we leverage recent results from the uncertain system simulation literature to derive an \emph{interval predictor} $[\underline{x}(t),\overline{x}(t)]$ for the system \eqref{eq:dynamics}, illustrated in \Cref{fig:prediction}. For any $N\in\Natural$, this predictor takes the information on the current state ${x}_N$, the confidence region $\cC_{[N],\delta}$, planned control sequence $\bu$ and admissible perturbation bounds $[\underline{\omega}(t),\overline{\omega}(t)]$; and must verify the \emph{inclusion property}:
\begin{equation}
\label{eq:inclusion-property}
\underline{x}(t)\leq x(t)\leq\overline{x}(t), \forall t\geq t_N.
\end{equation}

@@

\begin{figure}
	\begin{minipage}[b]{0.49\linewidth}
		\centering
		\includegraphics[trim={1cm 0 0 0}, clip, width=0.6\linewidth]{img/ellipsoid}
		\caption{The model estimation procedure, running on the obstacle avoidance problem of \Cref{sec:experiments}. The confidence region $C_{[N],\delta}$ shrinks with the number of samples $N$.}
		\label{fig:estimation}
	\end{minipage}
	\hfill
	\begin{minipage}[b]{0.49\linewidth}
		\centering
		\includegraphics[trim={4cm 0 0 0}, clip, width=0.8\linewidth]{img/obstacle_small}
		\caption{The state prediction procedure running on the obstacle avoidance problem of \Cref{sec:experiments}. At each time step (red to green), we bound the set of reachable states under model uncertainty \eqref{eq:confidence}}
		\label{fig:prediction}
	\end{minipage}%
\end{figure}

@@

\textbf{Linear dynamics and quadratic costs} ~~ Several approaches have been proposed for cumulative regret minimisation in the LQ problem. In the \emph{Optimism in the Face of Uncertainty} paradigm, the best possible dynamics within a high-confidence region is selected under a controllability constraint, to compute the corresponding optimal control in closed-form by solving a Riccati equation. The results of 	\citep{abbasi-yadkori11a,Ibrahimi2013,Faradonbeh2017} show that this procedure achieves a $\tilde{\cO}\left(N^{1/2}\right)$ regret. Posterior sampling algorithms \citep{Ouyang2017,abeille18a} select candidate dynamics randomly instead, and obtain the same result. Other works use noise injection for exploration such as \citep{Dean2017,Dean2018}. However, neither optimism nor random exploration fit a critical setting, where ensuring safety requires instead to consider pessimistic outcomes. The work of \citet{Dean2017} is close to our setting: after an offline estimation phase, they estimate a simple regret between a minimax controller and the optimal performance. Our work differs in that it addresses a generic shape cost. 
Another work of interest is \citep{Rosolia2019} where worst-case generic costs are considered. However, they assume the knowledge of the dynamics, and their rollout-based solution only produces inner-approximations and does not yield any guarantee. In this paper, interval prediction is used to produce oversets, while a near-optimal control is found using a tree-based planning procedure.

@@

We convert this confidence ellipsoid $\cC_{[N],\delta}$ from \eqref{eq:confidence-ellipsoid} into a polytope for $A(\theta)$. For simplicity, we present here a simple but coarse strategy: bound the ellipsoid by its enclosing axis-aligned hypercube:
\begin{align}
    \label{eq:polytope}
     A(\theta)\in \left\{ A_0 +\sum_{i=1}^{2^d}\alpha_{i}\Delta A_{i}: \alpha\geq 0,  \sum_{i=1}^{2^d}\alpha_{i}=1\right\} \text{ where } A_0 = A + \theta_{N,\lambda}^\transp\phi,\; h_i\in\{-1,1\}^d,
\end{align}
$\Delta A_{i} = {h_i} \sqrt{\frac{\beta_N(\delta)}{\lambda_{\max}(G_{N,\lambda})}}$. A tighter polytope derivation is presented in the Supplementary Material. % at the price of an increased computational cost required by the diagonalisation of $G_{N,\lambda}$.
